---
title: "Linear Regression Codes"
author: "Frankie de Jong"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8)
```

```{css, echo=FALSE}
h1, h4 {
  text-align: center;
  color: black;
  font-weight: bold;
}
```

## Section 1: Introduction to Regression/1.1: Baseball as a Motivating Example

```{r load lahman library,results='hide'}
library(Lahman)
```

### Filter Teams data frame to include years from 1961 to 2001 and Make Scatterplot of AB/G vs R/G

```{r warning=FALSE, message=FALSE}
library(tidyverse)
Teams %>% filter(yearID %in% 1961:2001) %>%
    mutate(AB_per_game = AB / G, R_per_game = R / G) %>%
    ggplot(aes(AB_per_game, R_per_game)) + 
    geom_point(alpha = 0.5)
```

### Scatterplot of W/G vs E/G

```{r}
Teams %>% filter(yearID %in% 1961:2001) %>%
    mutate(W_per_game = W / G, E_per_game = E / G) %>%
    ggplot(aes(W_per_game, E_per_game)) + 
    geom_point(alpha = 0.5)
```

### Scatterplot of XB2/G vs XB3/G

```{r}
Teams %>% filter(yearID %in% 1961:2001) %>%
    mutate(X2B_per_game = X2B / G, X3B_per_game = X3B / G) %>%
    ggplot(aes(X2B_per_game, X3B_per_game)) + 
    geom_point(alpha = 0.5)
```

## Section 1: Introduction to Regression/1.2: Correlation

### Correlation coefficient between AB and R per game

```{r}
Teams %>% filter(yearID %in% 1961:2001) %>%
    mutate(AB_per_game = AB / G, R_per_game = R / G) %>%
    summarize(r = cor(AB_per_game, R_per_game))
```

### Correlation coefficient between W and E

```{r}
Teams %>% filter(yearID %in% 1961:2001) %>%
          mutate(W_per_game = W / G, E_per_game = E / G) %>%
          summarize(r = cor(W_per_game, E_per_game))
```

### Correlation coefficient between X2B and X3B

```{r}
Teams %>% filter(yearID %in% 1961:2001) %>%
          mutate(X2B_per_game = X2B / G, X3B_per_game = X3B / G) %>%
          summarize(r = cor(X2B_per_game, X3B_per_game))
```

## Section 1: Introduction to Regression/1.3: Stratification and Variance Explained

### Define female_heights
```{r}
set.seed(1989) #if you are using R 3.5 or earlier
set.seed(1989, sample.kind="Rounding") #if you are using R 3.6 or later
library(HistData)
data("GaltonFamilies")

female_heights <- GaltonFamilies%>%     
    filter(gender == "female") %>%     
    group_by(family) %>%     
    sample_n(1) %>%     
    ungroup() %>%     
    select(mother, childHeight) %>%     
    rename(daughter = childHeight)
```

### Calculate mean heights, sd, cor

```{r}
mu_mom <- mean(female_heights$mother)
sd_mom <- sd(female_heights$mother)
mu_dau <- mean(female_heights$daughter)
sd_dau <- sd(female_heights$daughter)
cc <- cor(female_heights$daughter, female_heights$mother)
m <- cc*sd_dau/sd_mom  
b <- mu_dau - m*mu_mom
(m*2+b) - (m*1+b)
cc*cc*100
x = 60
m*x+b
```

